test

### 2020-05-25 Action letter.Rmd

\WorkInProgress

@@

> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers' qualitative involvement [@flouriFathersBehaviorsChildren2010]. Although some research suggests that both parents influence children similarly [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014], there is also evidence for unique contributions of mothers and fathers. Some have theorized that mothers and fathers might fulfill somewhat specialized roles [@bogelsDoesFatherKnow2011]. In line with the theoretical notions of "attachment" and "activation" relationships, mothers tend to encourage and respond supportively to emotional expression to a greater extent than fathers, whereas fathers structure and sanction emotion expressions more than mothers [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. Moreover, for adolescents' emotional adjustment, maternal warmth was found to be a more salient socialization mechanism than demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. Similar differentiated associations were also found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is some evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is incongruous with the notion of the maternal "attachment" relationship. We might therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demanding behaviors, such as consistency and inductive reasoning, $H_{3}$.

@@

> We found only limited evidence for within-family predictive effects, most of which were child effects on parents. Instead, we mostly found evidence for between-family level correlations. We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016]. In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a]. The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age. They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age. The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.  
  As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory. Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time. Our findings serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general. We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?  
  Future research might also devote greater attention to between-family level correlations as an outcome of interest. These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs. We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors. For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990].

>>>>

We have done as requested in the re-written section on *Differences between mothers and fathers*:

> The notion that fathers should share child-rearing responsibilities with mothers has become widely accepted in the Western world [@pleckFatherhoodIdealsUnited1997]. Consequently, researchers have come to study qualitative involvement along the same dimensions for mothers and fathers [@maccobySocializationContextFamily1983; @pleckPaternalInvolvementRevised2004]. Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010]. Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014]. Others have argued that mothers and fathers might contribute through somewhat different mechanisms. Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].  
  In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers. Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.

@@

> Fathers' qualitative involvement is associated with emotional adjustment, even after accounting for mothers [@flouriFathersBehaviorsChildren2010]. Some research suggests that both parents influence children similarly, but that effects might be stronger for mothers [@lambDevelopmentSignificanceFatherChild2004; @faganShouldResearchersConceptualize2014]. Others have argued that mothers and fathers might contribute through somewhat different mechanisms. Attachment relationships are grounded in a balance between warmth and security, and exploration [@grossmannDeveloppementAttachementAdaptation1998]. Some have proposed that warmth is more central to mother-child attachment relationships, whereas father-child relationships can be construed as "activation relationships", with greater emphasis on exploration [@paquetteFatherChildActivationRelationship2013]. By encouraging children to venture outside their comfort zone, fathers stimulate a broad spectrum of emotions, thereby expanding children's emotional repertoire, and engaging their developing regulatory abilities  [@flouriFathersBehaviorsChildren2010; @bogelsDoesFatherKnow2011; @parkeFatherInvolvement2000].  
  In line with this theory, empirical research has found that mothers' responses to emotional expression are more warm and supportive than fathers' responses, which are more demanding and punitive [@cassanoInfluenceGenderParental2007; @eisenbergParentalReactionsChildren1999]. In adolescence, maternal warmth was found to be a more salient socialization mechanism of emotional adjustment than maternal demandingness, whereas this pattern was reversed for fathers [@vanlissaRoleFathersMothers2018a]. Similar differentiated associations were found for younger children's psychological adjustment [@mcdowellDifferencesMothersFathers2003]. Finally, there is initial evidence that emotional adjustment is more negatively impacted by mothers' negative parenting, compared to fathers [@changHarshParentingRelation2003]; perhaps because negative parenting is somewhat incongruous with the warm "attachment" relationship typically expected of mothers. Based on this research, we therefore hypothesize that emotional adjustment would be more strongly associated with maternal warmth and anger, and with paternal demandingness, operationalized as consistency and inductive reasoning, $H_{3}$.

@@

> We found only limited evidence for within-family predictive effects, most of which were child effects on parents. Instead, we mostly found evidence for between-family level correlations. We are not the first to find that parent effects are less prevalent once stable between-family differences are accounted for [@keijsersParentalMonitoringAdolescent2016]. In the area of parenting and emotional adjustment development, we are aware of only one prior study that controlled for stable between-family differences, and similarly reported mostly child effects and between-family level correlations in a sample of adolescents [@vanlissaRoleFathersMothers2018a]. The authors explained their findings by proposing that child effects might be more common in adolescence than in childhood, because parent-adolescent relationships are more egalitarian at that age. They argued that the between-family level correlations might reflect the consolidated influence of within-family parent effects that occurred at an earlier age. The fact that the present study similarly found a preponderance of child effects and between-family level correlations in a sample of younger children, age 6-12, suggests that either parent effects occurred at an even younger age, or that something else is at play.  
  As similar patterns continue to emerge, it might be worth reflecting on their implications for developmental theory. Our findings are less congruent with a fully bidirectional conceptualization of reciprocal influence, and more consistent with a model where parents "set the stage" for children's development through their stable differences from other parents, and continue to adjust to children's changing needs over time. Our findings serve as a reminder that we should be critical of causal interpretations of research that does not control for stable between-family differences, not only in the area of emotional adjustment [e.g., @brandPerceivedParentalInconsistency1990], but in developmental research in general. We should reconsider the question: To what extent are parents shaping children's development, and to what extent are they adapting to it?  
  Future research might also devote greater attention to between-family level correlations as an outcome of interest. These correlations can reflect the consolidated effects of parenting effects that occur at an early age [@vanlissaRoleFathersMothers2018a], or alternatively, they can reflect the influence of third variables that impact both constructs. We should seek to understand why between-family level correlations are so prominent; for example, by comparing their strength in samples with children of different ages, including very young children, or by examining the influence of more sociological factors. For example, economic hardship undermines parents' capacity for consistent parenting, as well as children's socio-emotional adjustment [@mcloydImpactEconomicHardship1990]. Due to disciplinary differences, such sociological factors are rarely examined in conjunction with sophisticated models of reciprocal influences between parents and children.

### 2020-05-25 Paper_shortened_pdf.Rmd

To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-individual differences (the mean level for that individual across all measurement occasions), and within-individual fluctuations (deviations from this person-specific mean). The ICCs (supplementary Table S1) indicate the proportion of variance at the between- versus within-family level. For warmth, consistency, and anger, the ICCs were high, ranging from [.60, .64]. This indicates that about 60% of the variance in these parenting behaviors is due to between-family differences. For inductive reasoning, they were notably lower, [.46, .50]. For emotional adjustment, they were even lower, $r = .25$.

@@

table_within_cor <- lapply(cor_tables, `[[`, "Within.Std.Averaged.Over.Clusters")
table_within_cor <- order_by_name(table_within_cor)
table_within_cor <- do.call(cbind, table_within_cor)
table_within_cor <- cbind(rename_pars(rownames(table_within_cor)), table_within_cor)
table_within_cor <- as.matrix(table_within_cor)
colnames(table_within_cor) <- c("Parameter", rep(table_within_cor[,1], 4))
rownames(table_within_cor) <- NULL
table_within_cor <- gsub("\\*+", "*", table_within_cor)
table_within_cor <- gsub("^1$", "", table_within_cor)
table_within_cor <- table_within_cor[-3, -seq(2, 11, by = 3)]
table_within_cor <- rbind(c("Within-families level correlations", rep(NA, ncol(table_within_cor)-1)), table_within_cor)

@@

tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
names(tab)[1] <- "Param."
reg_table(tab, caption = "Standardized between-family level intercepts (average effect) and unstandardized variances of within-family predictive effects.", note = "Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate. Matching letters indicate that parameters differed significantly; see Supplementary Table S4 for all difference tests.")
```

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on random intercepts of:",
                              "Effects of relative quantitative involvement on random intercepts of:")
tabinv$Parameter <- gsub("^(.+) on (?!random).*$", "\\1", tabinv$Parameter, perl= TRUE)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on parenting effects of:",
                              "Moderating effects of relative involvement on parenting effects of:",
                              "Moderating effects of joint involvement on child effects on:",
                              "Moderating effects of relative involvement on child effects on:")
tab <- rbind(tabinv, tabinvmod)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
names(tab)[1] <- "Param."
cat(as.character(reg_table(tab, caption = "Standardized regression coefficients of quantitative involvement on between-family differences in qualitative parenting and emotional adjustment, and on within-family predictive effects", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")))
```

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Sex differences in random intercepts of parenting and adjustment",
                              "Moderation of parenting effects by sex",
                              "Moderation of child effects by sex")
#tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
names(tab)[1] <- "Param."
reg_table(tabsex, caption = "Standardized regression coefficients of child sex on average levels of qualitative parenting and emotional adjustment, and on within-family predictive effects.", note = "Child sex is a binary predictor (-1 = female, 1 = male). The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

>>>>

To address our research questions, we used multilevel, autoregressive structural equation modeling [@asparouhovDynamicStructuralEquation2018]. In this approach, the variance in qualitative parenting behaviors and emotional adjustment is partialized into stable between-family differences (the mean level for each family across all measurement occasions), and within-family fluctuations (within-family deviations from the family-specific mean). The ICCs (supplementary Table S1) indicate the proportion of variance at the between- versus within-family level. For warmth, consistency, and anger, the ICCs were high, ranging from [.60, .64]. This indicates that about 60% of the variance in these parenting behaviors is due to between-family differences. For inductive reasoning, they were notably lower, [.46, .50]. For emotional adjustment, they were even lower, $r = .25$.

@@

table_within_cor <- lapply(cor_tables, `[[`, "Within.Std.Averaged.Over.Clusters")
table_within_cor <- order_by_name(table_within_cor)
table_within_cor <- do.call(cbind, table_within_cor)
table_within_cor <- cbind(rename_pars(rownames(table_within_cor)), table_within_cor)
table_within_cor <- as.matrix(table_within_cor)
colnames(table_within_cor) <- c("Parameter", rep(table_within_cor[,1], 4))
rownames(table_within_cor) <- NULL
table_within_cor <- gsub("\\*+", "*", table_within_cor)
table_within_cor <- gsub("^1$", "", table_within_cor)
table_within_cor <- table_within_cor[-3, -seq(2, 11, by = 3)]
table_within_cor <- rbind(c("Within-family level correlations", rep(NA, ncol(table_within_cor)-1)), table_within_cor)

@@

tab$Parameter <- gsub("^Variance", "Unstandardized variance", tab$Parameter)
names(tab)[1] <- "Param."
reg_table(tab, caption = "Average of the standardized within-family effects and variances of unstandardized within-family effects.", note = "Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate. Matching letters indicate that parameters differed significantly; see Supplementary Table S4 for all difference tests.")
```

@@

```{r tabinvolvement, results = "asis"}
tabinv <- tab_std_sections$`Effects of quantitative involvement on parenting and adjustment`[c(1:4, 1, 5:7),]
tabinv$Parameter[c(1,5)] <- c("Effects of joint quantitative involvement on between-family differences in:",
                              "Effects of relative quantitative involvement on between-family differences in:")
tabinv$Parameter <- gsub("^(.+) on (?!random).*$", "\\1", tabinv$Parameter, perl= TRUE)
#reg_table(tabinv, caption = "Effects of quantitative involvement on parenting and adjustment", note = "Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
tabinvmod <- rbind(
  tab_std_sections$`Moderating effect of quantitative involvement on parenting effects`[c(1:3, 1, 4:5),],
  tab_std_sections$`Moderating effect of quantitative involvement on child effects`[c(1:3, 1, 4:5),])
tabinvmod$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabinvmod$Parameter)
tabinvmod$Parameter[c(1,4,7,10)] <- c("Moderating effects of joint involvement on within-family parenting effects",
                              "Moderating effects of relative involvement on within-family parenting effects",
                              "Moderating effects of joint involvement on within-family child effects",
                              "Moderating effects of relative involvement on within-family child effects")
tab <- rbind(tabinv, tabinvmod)
#tab[is.na(tab)] <- ""
rownames(tab) <- NULL
names(tab)[1] <- "Param."
cat(as.character(reg_table(tab, caption = "Standardized effects of joint and relative quantitative involvement.", note = "The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")))
```

@@

```{r tabsex, results = "asis"}
tabsex <- rbind(
  tab_std_sections$`Sex differences in parenting and adjustment`,
  tab_std_sections$`Sex differences in cross-lagged effects`[c(1:3, 1, 4:5),])
tabsex$Parameter <- gsub("^.+?(Mother|Father).+?$", "\\1", tabsex$Parameter)
tabsex$Parameter <- gsub(" on sex$", "", tabsex$Parameter)
tabsex$Parameter[c(1,5,8)] <- c("Effect of sex on between-family differences in of parenting and adjustment",
                              "Moderating effects of sex on within-family parenting effects",
                              "Moderating effects of sex on within-family child effects")
#tabsex[is.na(tabsex)] <- ""
rownames(tabsex) <- NULL
names(tab)[1] <- "Param."
reg_table(tabsex, caption = "Standardized effects of child sex.", note = "Child sex is a binary predictor (-1 = female, 1 = male). The effects reported in this Table correspond to the 'Between-family level' panel of Figure 1. Param.: Parameter. Est.: Estimate. HDPI: Highest posterior density interval. Emo.: Emotional adjustment. *: <5% of posterior has opposite sign from estimate.")
```

### 2020-05-25 Revision_2.Rmd

@@

\vspace{\baselineskip}
Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Ample research has demonstrated associations between children's emotional adjustment and parental involvement; both quantitative involvement (the amount of time parents spend caring for their child), and qualitative involvement (specific parenting behaviors). Nevertheless, the existing literature has two major shortcomings. Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together, let alone in interaction [@craigDoesFatherCare2006; @faganShouldResearchersConceptualize2014]. The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

@@

Early fatherhood research primarily operationalized involvement in *quantitative* terms: Father absence in divorced families, and childcare sharing in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011]. The equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001]. Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigDoesFatherCare2006]. This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].

@@

It has been argued that within-family predictive effects are of prime theoretical relevance in developmental psychology [@keijsersParentalMonitoringAdolescent2016]. And yet, our study and others [e.g. @vanlissaRoleFathersMothers2019] have found high intraclass correlations, meaning that the brunt of variance in the variables studied is due to stable between-family differences. This suggests that between-family level correlations should be treated as an outcome of interest.

Future research could seek to understand why evidence for between-family level correlations outweighs evidence for within-family parenting effects. We argued elsewhere that between-family level correlations can have different causes [@vanlissaRoleFathersMothers2019]. They might, for example, reflect the consolidated effect of within-family parental socialization at an earlier age. This could be tested by comparing the size of between-family correlations across children of different ages, including very young children. Alternatively, these correlations might reflect the influence of stable third variables that impact both constructs: Economic hardship, for instance, undermines both parents’ capacity for consistent parenting, and children’s socio-emotional adjustment [@mcloydImpactEconomicHardship1990]. Testing this explanation would require extensive investigation of confounding variables. This is common in sociology, but has rarely been done in conjunction with sophisticated models of development [see @keizerRoleFathersIntergenerational2020]. Finally, future research might shed light on the extent to which between-family level correlations reflect genetic co-variation, or shared- versus non-shared environmental effects by using genetically informed designs with twin-families or adoptive versus biological families [see @knopikBehavioralGenetics2017].

>>>>

authornote: |
  All code and output are available from the OSF (https://osf.io/678rq/).
  The first author is supported by a NWO Veni grant (NWO
  grant number VI.Veni.191G.090). The second author is supported by a NWO Vidi
  Grant (NWO grant number 452-17-006).

@@

\vspace{\baselineskip}
Emotional adjustment is defined as the absence of internalizing symptoms and emotional dysregulation [@goodmanPsychometricPropertiesStrengths2001]. It is integral to social competence, academic achievement, and mental health [@grossHandbookEmotionRegulation2013]. Ample research has demonstrated associations between children's emotional adjustment and parental involvement; both quantitative involvement (the amount of time parents spend caring for their child), and qualitative involvement (specific parenting behaviors). Nevertheless, the existing literature has two major shortcomings. Firstly, although it is widely acknowledged that quantitative and qualitative involvement are both relevant [@faganShouldResearchersConceptualize2014; @pleckPaternalInvolvementRevised2004], these dimensions are rarely investigated together, let alone in interaction [@craigHowMothersFathers2011; @faganShouldResearchersConceptualize2014]. The present study overcomes this limitation by investigating both qualitative and quantitative involvement, and the interplay between them, in relation to children's emotional adjustment.

@@

Early fatherhood research primarily operationalized involvement in *quantitative* terms: Father absence in divorced families, and childcare sharing in dual-earner families [@lambBiosocialPerspectivePaternal1987]. Quantitative involvement has been researched extensively in terms of the absolute amount of childcare provided [see @pleckPaternalInvolvementRevised2004]. However, some have argued that the *division* of childcare tasks is more relevant for children's adjustment [e.g., @keizerInfluenceFathersMothers2019; @craigHowMothersFathers2011]. The equity of the childcare distribution is associated with socio-emotional development [@deutschPaternalParticipationChild2001]. Moreover, an unequal distribution often implies that women perform a disproportionate share of the work [@craigHowMothersFathers2011]. This strains the marital relationship [@shockleyItNotWhat2018], and erodes parenting quality, particularly for parents of difficult children [@sheltonInterparentalConflictNegative2008].

@@

It has been argued that within-family predictive effects are of prime theoretical relevance in developmental psychology [@keijsersParentalMonitoringAdolescent2016]. And yet, our study and others [e.g. @vanlissaRoleFathersMothers2019] have found high intraclass correlations, meaning that the brunt of variance in the variables studied is due to stable between-family differences. These data suggest that between-family level correlations should be treated as an outcome of interest. We argued elsewhere that between-family level correlations can have different causes [@vanlissaRoleFathersMothers2019]. They might reflect the consolidated effect of within-family parental socialization at an earlier age. This could be tested by comparing the size of between-family correlations across children of different ages, including very young children. Alternatively, they might reflect the influence of third variables that impact both constructs: Economic hardship, for instance, undermines both parents’ capacity for consistent parenting, and children’s socio-emotional adjustment [@mcloydImpactEconomicHardship1990]. Testing this explanation would require extensive investigation of confounding variables. This is common in sociology, but has rarely been done in conjunction with sophisticated models of development [see @keizerRoleFathersIntergenerational2020]. Finally, future research might shed light on the extent to which between-family level correlations reflect genetic co-variation, or shared- versus non-shared environmental effects, by using genetically informed designs with twin- or adoptive families [see @knopikBehavioralGenetics2017].

### 2020-05-25 manuscript.Rmd

authornote: |
  All code and output are available from the OSF (https://osf.io/678rq/).
  The first author is supported by a NWO Veni grant (NWO
  grant number VI.Veni.191G.090). The second author is supported by a NWO Vidi
  Grant (NWO grant number 452-17-006).

@@

Finally, negative parenting, which we operationalized as parental anger, is a qualitative parenting behavior negatively assosicated with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003]. We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.

@@

There are also methodological advantages to focusing on the distribution of childcare. Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving *more* total care [see @deutschPaternalParticipationChild2001]. This confound can be mitigated by estimating a latent mean and difference [@mcardleLatentVariableModeling2009]. The latent mean reflects joint involvement; the total (average) amount of care provided to children, and tge latent difference reflects relative involvement; the extent to which the mother is more involved than the father (i.e., the childcare distribution).

@@

Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation; the tables and figure were created using tidySEM. All code and output are available online at <https://osf.io/678rq/>. DSEM can only be estimated using a Bayesian estimator. This is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parametric MCAR test resulted in $p > .05$). We used two Gibbs-sampler chains of 50000 iterations each, with normal priors centered on zero for regression coefficients, and mildly informative inverse Wishart priors for covariance matrices. We assessed convergence through visual inspection of the chains and posterior distributions. We report the posterior median, and a 95% highest posterior density interval (HPDI). This is the "narrowest" possible interval that contains 95% of samples from the posterior distribution. It represents the most likely population values of the parameter in question. Results are best interpreted by examining the effect size (correlations and standardized regression coefficients), and the width of the HPDI. Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate. Supplementary Table S4 reports all standardized and unstandardized model parameters.

>>>>

authornote: |
  All code, output, and supplemental materials are available on the OSF
  (https://osf.io/678rq/).
  The present study was supported by grants from the Netherlands Organization
  for Scientific Research (NWO MaGW VIDI; grant no. 452-17-005 to R.K.; NWO
  SGW VENI; grant no. VI.VENI.191G.090 to C.J.), and by a grant from the 
  European Research Council (ERC StG; grant no. 757210 to R.K.).

@@

Finally, negative parenting, which we operationalized as parental anger, is a qualitative parenting behavior negatively associated with children's emotional adjustment [@changHarshParentingRelation2003; @eisenbergParentalReactionsChildren1999]. Parental anger negatively impacts children's emotional adjustment [@scaramellaParentingChildrenDistress2008], because it is upsetting, and because anger reflects poor parental emotion regulation, which serves as a model for children [@changHarshParentingRelation2003]. We thus hypothesized that anger would be negatively related to emotional adjustment, $H_{1d}$.

@@

There are also methodological advantages to focusing on the distribution of childcare. Operationalizing quantitative involvement in terms of both parents' absolute involvement incurs a potential confound: Effects that appear to be due to fathers' greater involvement might be due to children receiving *more* total care [see @deutschPaternalParticipationChild2001]. This confound can be mitigated by estimating a latent mean and difference [@mcardleLatentVariableModeling2009]. The latent mean reflects joint involvement; the total (average) amount of care provided to children, and the latent difference reflects relative involvement; the extent to which the mother is more involved than the father (i.e., the childcare distribution).

@@

Analyses were conducted using Mplus 8.0 and the R-package MplusAutomation; the tables and figure were created using tidySEM. All code, output, and supplemental materials are available on the OSF (https://osf.io/678rq/). DSEM can only be estimated using a Bayesian estimator. This is a full-information estimator, which means that all available information is used without imputing missing values, and consistent estimation is guaranteed under the assumption that data are missing-at-random (a non-parametric MCAR test resulted in $p > .05$). We used two Gibbs-sampler chains of 50000 iterations each, with normal priors centered on zero for regression coefficients, and mildly informative inverse Wishart priors for covariance matrices. We assessed convergence through visual inspection of the chains and posterior distributions. We report the posterior median, and a 95% highest posterior density interval (HPDI). This is the "narrowest" possible interval that contains 95% of samples from the posterior distribution. It represents the most likely population values of the parameter in question. Results are best interpreted by examining the effect size (correlations and standardized regression coefficients), and the width of the HPDI. Significance is not a Bayesian concept, but we use the shorthand "significant" to refer to estimates for which less than 5% of the posterior distribution had the opposite sign from the estimate. Supplementary Table S4 reports all standardized and unstandardized model parameters.
